---
title: "MI Economic Indicator Report"
author: "Justin Baker"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_notebook: default
---
## Reporting Time!
Now let's take everything we've learned and generate a report. But before we begin, a few notes.

#### A Note About Style
You may notice some differences between the way Alicia, Ryan and I have written some things in our code. There are many ways to do many things in R, so naturally everyone's code may end up looking a little different.

One way to make sure everyone's code isn't too different is to implement a programming style guide in your organization. In the Office of Good Government at  the State of Michigan, we use [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml), with a few modifications. That helps make sure everyone can read everyone else's code.

#### Do It Yourself!
I've used some data from `future.michigan.gov`, our open performance site. But I strongly encourage you to visit your own Open Data Portal and find some data that you know well to build charts of your own!

Okay, back to the report!

## Packages
Before we begin, let's make sure we've installed all the packages we'll need. If you haven't installed the following packages, head over to the console and run `install.packages("package_name")`.


```{r setup, warning=FALSE, message=FALSE}
#### Libraries ####
# Libraries
library(flexdashboard)
library(tidyverse)
library(RSocrata)
library(readxl)
library(lubridate)

#### Options ####
options(tz="US/Michigan")  # might be req for read.socrata()
knitr::opts_chunk$set(fig.width=16, fig.height=8)

```

## Unemployment Data
Next, we'll practice what Alicia and Ryan have just shown us - we'll pull, clean, and transform datasets from Socrata and Excel to create the data that will generate our report.

Notice that our Socrata data is pulled through the Socrata API. This means that if I want to update this report in the future, all I need to do is re-run the code chunks, and all of my visualizations will automatically update.

```{r unemployment_data, include=FALSE}

#### MI Economic Outcomes ####

# Unemployment, Labor Force, LFPR, U-6
econ_url <- "https://future.michigan.gov/Economy/LMISI-Monthly-Measures/v2fn-bucf"
econ <- read.socrata(econ_url)


#### Join MI Unemployment Data ####

# Unemployment Rates
unemp_url <- "https://data.michigan.gov/Economy/Unemployment-Rate/kamp-ngb9"
unemp <- read.socrata(unemp_url) %>%
  # convert to tibble class
  as_tibble() %>%
  # remove unwanted rows
  select(-highlights, -rank) %>%
  # reformat rates as percentages
  mutate(unemployment_rate_us = unemployment_rate_us * 100,
         unemployment_rate_mi = unemployment_rate_mi * 100,
  # reformat dates
         date = as_date(parse_date_time(date, orders = "m-y"))) %>%
  # rename columns
  rename(MI = unemployment_rate_mi,
         US = unemployment_rate_us)


#### FRED County Data ####
counties <- read_csv("fredgraph.csv") %>%
  rename(Wayne = MIWAYN3URN,
         Kalamazoo = MIKALA7URN,
         Saginaw = MISAGI5URN,
         Ingham = MIINGH5URN)


#### Join Unemployment Data ####

unemployment <- unemp %>%
  # return matching rows from counties
  left_join(counties, by = c("date" = "DATE"))

```


```{r chart_theme, include=FALSE}

# Please feel free to use this theme in your visualizations! I've just made a few modifications to the theme_minimalist() ggplot theme to remove axis titles. We'll see that come in handy in the dashboards section!

# ggplot theme specifications
theme_som <- function(base_size = 11, 
                      base_family = "",
                      base_line_size = base_size / 22,
                      base_rect_size = base_size / 22) {
  # Scaled down theme_minimalist
  theme_bw(
    base_size = base_size,
    base_family = base_family,
    base_line_size = base_line_size,
    base_rect_size = base_rect_size
  ) %+replace%
    theme(
      axis.ticks        = element_blank(),
      legend.background = element_blank(),
      legend.key        = element_blank(),
      panel.background  = element_blank(),
      panel.border      = element_blank(),
      strip.background  = element_blank(),
      plot.background   = element_blank(),
      axis.title        = element_blank(),

      complete = TRUE
    )
}
```

******

## Michigan Economic Indicators

### Unemployment

```{r unemployment_plot}

unemployment %>%
  # select unemp columns
  select(date, MI, US) %>%
  # the data is wide - let's make it tidy!
  gather(key = "Location", value = "unemp_rate", -date) %>%
  # plotting our data
  ggplot(aes(date, unemp_rate, color = Location)) +
    geom_line() +
    theme_som()

```

### Labor Underutilization

```{r u-6}
econ %>%
  # U-6 has some NA - let's get rid of those
  filter(!is.na(u_6)) %>%
  ggplot(aes(date, u_6)) +
    geom_line() +
    theme_som()
```

### Labor Force

```{r lf}
econ %>%
  ggplot(aes(date, labor_force)) +
  geom_line() +
  theme_som()
```

### Labor Force Participation Rate

```{r lfpr}
econ %>%
  ggplot(aes(date, labor_force_participation_rate)) +
  geom_line() +
  theme_som()
```
