---
title: "Read Data"
output: html_document
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,warning = FALSE)
library(dplyr)
library(httr)
library(jsonlite)
library(kableExtra)
library(knitr)
library(readr)
library(RSocrata)
library(lubridate)
```

## RSocrata

[RSocrata](https://github.com/Chicago/RSocrata) is a library created by the City of Chicago to enable reading, writing and discovering data on Socrata powered sites.

- Read a Socrata dataset into R as a data frame.
- We can use the full human-readbale URL or the SODA API endpoint. 

### Fish Species in Michigan

- Source: https://data.michigan.gov/Environment/Michigan-Fish/he9h-7fpa
- Since this is a public dataset, we don't need to pass user name, password or app token
- We can even display the image URL as an image

Run the chunk below to take a look at the dataset!

```{r rSocrataExample}
fish_url <- "https://michigan-scgc.data.socrata.com/Environment/Michigan-Fish/he9h-7fpa"
rsocrata_data <- read.socrata(fish_url)

rsocrata_data %>%
  top_n(10, commonname) %>%
  mutate(Image = paste0("<image src='",imageurl,"'></image>")) %>%
  # setting escape to FALSE allows html to be rendered for the fish images!
  kable(escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(width = "900px", height = "300px")
```

### Local Area Unemployment Statistics (LAUS) from the State of California Employment Development Department

Let's read in data on unemployment rates, coming to us for this example from California.

- Source: https://data.edd.ca.gov/Labor-Force-and-Unemployment-Rates/Local-Area-Unemployment-Statistics-LAUS-/e6gw-gvii
- This is Open Data and requires no username/password to access. 

```{r read LAUS}
edd_url <- "https://data.edd.ca.gov/Labor-Force-and-Unemployment-Rates/Local-Area-Unemployment-Statistics-LAUS-/e6gw-gvii"
laus_data <- read.socrata(edd_url)
str(laus_data)
```

### Use RSocrata::read.socrata() on your own! 
Find a dataset on a Socrata site, like https://www.dallasopendata.com/, and use RSocrata::read.socrata() to read it in.

```{r, eval = F}
url <- "https://data.michigan.gov/Economy/Michigan-Per-Capita-Personal-Income/7frj-2rv6"
mi_cap_data <- read.socrata(url)
summary(mi_cap_data)
```

### Read Private Data with RSocrata: Example with Stellar Sea Lion Count

- For private datasets, you do need to include user name and password.
- You can also create a [Socrata API Key](https://socrataapikeys.docs.apiary.io/#) and pass the Key Id and Key Secret as user name and password, respectively

***Not really a private dataset, but made private for this example***

```{r SocrataExampleAuth}
url <- "https://alicia.data.socrata.com/Species/2016-Counts-of-Steller-Sea-Lions-Pups-Adults-and-J/qaj3-k28x"

email <- Sys.getenv("MY_SOCRATA_USERNAME")
password <- Sys.getenv("MY_SOCRATA_PASSWORD")
rsocrata_data <- read.socrata(url, app_token = NULL, email = email, password = password, stringsAsFactors = TRUE)

rsocrata_data %>%
  top_n(10, Site.Name) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(width = "900px", height = "300px")
```


## Readr package to read datasets from Socrata

- Can read public datasets using the CSV endpoint (which returns all of the data! No need to paginate through a RESTful API)
- Preserves field names as-is (rather than replacing spaces with periods)
- Fast!
- Returns as Tibble (part of the tidyverse!)

### Here's an Example: Unemployment Insurance Weekly Claims Data for California
With readr, you can rely on it to reasonably interpret column data types, or you can set them yourself in a few ways.

```{r}
# public dataset example requiring no authentication
uiwc_data <- read_csv("https://data.edd.ca.gov/api/views/qbrj-7mnz/rows.csv?accessType=DOWNLOAD&bom=true&format=true", 
                      col_types = cols(
                        `Area Type` = "c",
                        `Area Name` = "c",
                        `Filed week ended` = col_date(format = "%m/%d/%Y"),
                        `Initial Claims` = "n",
                        `Reflecting Week Ended` = col_date(format = "%m/%d/%Y"),
                        `Continued Claims` = "n",
                        `Covered Employment` = "n",
                        `Insured Unemployment Rate` = "d"
                      ))
str(uiwc_data)
```

## Try using readr with the CSV endpoint!

First up, use readr with the Socrata dataset you chose earlier (the one you used with RSocrata)
1. Right click on the CSV or CSV for Excel button on the Socrata dataset page and copy that link
2. Use readr::read_csv() on that endpoint, like above (Don't worry about the col_types bit, you can leave that out for now).

```{r CSVEndpoint,eval=F}

```


## httr + jsonlite + SODA2

Local Area Unemployment Statistics from the JSON endpoint of SODA, Socrata Open Data API
- R supports all HTTP requests methods
- JSONLite makes it easy to work with JSON source
- Use the SODA2 endpoint (which defaults to 1,000 records, but can get rowcount and then set limit to it)

```{r httr LAUS}
laus_soda2_json <- "https://data.edd.ca.gov/resource/e6gw-gvii.json?$limit=134000"

response <- GET(url = laus_soda2_json) %>%   # << make's the request to the Socrata API endpoint
            content("text") %>%              # << reads the byte response as text
            fromJSON()                       # << interprets the JSON structure returned in the response as a data frame
str(response)
```

## Try it yourself: Use httr + jsonlite + SODA2
With the Socrata dataset you picked earlier: 
- copy the SODA json endpoint (from the API menu)
- optional: add a limit parameter (like the example above) to grab as many rows as you need (be kind to the hotel wifi)
- use httr::GET(), content(), and jsonlite::fromJSON() to read the data

```{r httrPractice, eval=F}

```

Now, get ready to Visualize some of this data in the next section!
