---
title: "Visualize Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(RSocrata)
library(dplyr)
library(ggplot2)
library(gganimate)
library(knitr)
library(kableExtra)
```

## Visualize with ggplot2

ggplot2 is a package for visualizing data in R. Think of it as a "grammar of graphics".
With ggplot2, one can create all manner of visualizations. We we also explore gganimate, a package for bringing a visual to life. 

## Read in some Steller Sea Lion counts from NOAA Alaska Open Data

```{r steller sea lions data}
sea_lions_url <- "https://noaa-fisheries-afsc.data.socrata.com/Species/2016-Counts-of-Steller-Sea-Lions-Pups-Adults-and-J/cntk-5zqj"
rsocrata_data <- read.socrata(sea_lions_url)

cat("Let's take a look at a subset of the rows.\n")
head(rsocrata_data, n = 5) %>%  kable(format = 'markdown', align = "c")
```

## And Group by Region

```{r region}
region_counts <- group_by(rsocrata_data, Region) %>% 
  summarise(Non.Pup = sum(Non.Pup, na.rm = TRUE),
            Pup = sum(Pup, na.rm = TRUE))
region_counts
```

## Visually Display the Number of Pups and Non-Pups by Site

The Grammar of Graphics offers up *three* key elements when creating a visualization:
* data, or the dataset to be used in the visualization
* geom, or what type of visualization will be created
* aes, or the details of mapping the geom to the data, plus aesthetics

The structure:

ggplot( *data to be used* ) +
  *geom to be used* ( mapping = *aes( like what is x, y, etc)* )

```{r seals chart, echo=TRUE}
ggplot(data = region_counts) +
  geom_col(mapping = aes(x = Region, y = Non.Pup))
```

## Try it on your own:
Create the same type of chart, but showing the count of *Pups* on the y-axis. 

```{r try seals chart, eval=F}


```

## Trees! (in Edmonton)

Reading in the Trees from Edmonton's Open Data site ()

```{r edmonton trees}
trees_url <- "https://data.edmonton.ca/resource/93cp-z7sw.json?$limit=100000"  ## Just 100,000 records to make things a little faster
trees <- read.socrata(trees_url) %>% mutate(condition_percent = as.numeric(condition_percent),
                                            diameter_breast_height = as.numeric(diameter_breast_height))
```

### Trees but with readr

```{r readr trees}
readr_trees <- read_csv("https://data.edmonton.ca/api/views/eecg-fc54/rows.csv?accessType=DOWNLOAD")
str(readr_trees)

```


## Tree Size and Condition Percent by Genus
Here, we will use two additional arguments in our geom: 
*color* to visually distinguish Genus
*alpha* to control transparency of the points
*position* to modify how the points are plotted

```{r trees chart}
ggplot(trees) +
  geom_point(mapping = aes(x = diameter_breast_height, y = condition_percent, color = genus, alpha = 0.8), position = "jitter")
```

In the code chunk below, create the same chart but with colors representing 'location_type'. Some structure has been provided, but not all. 

```{r try trees chart,eval=F}
ggplot(trees) +
  geom_point(mapping = aes(), position = "jitter")
```

## Let's make a histogram with the trunk size of each tree

```{r histogram trees}
ggplot(readr_trees) +
  geom_histogram(aes(DIAMETER_BREAST_HEIGHT), binwidth = 5, color="white") +
  theme_linedraw()
```



## Faceting

```{r facet histogram}
ggplot(readr_trees) +
  geom_histogram(aes(DIAMETER_BREAST_HEIGHT), binwidth = 5, color="white") +
  facet_wrap( ~ GENUS) +
  theme_dark()
```


## Local Area Unemployment Statistics for Los Angeles
### Visualize labor force, unemployment rate for areas in California

```{r LAUS exploration LA}
laus_la <- laus_data %>% filter(`area_name` == "Los Angeles County" & `seasonally_adjusted_y_n` == "Y")
ggplot(laus_la) +
  geom_line(mapping = aes(x = date, y = `unemployment_rate`)) 

```

```{r LAUS MSA}
laus_msa <- laus_data %>% filter(`area_type` == "Metropolitan Area" & `seasonally_adjusted_y_n` == "Y")
ggplot(laus_msa) +
  geom_line(mapping = aes(x = date, y = `unemployment_rate`,color = `area_name`))

```


















