---
title: "Visualize Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(RSocrata)
library(dplyr)
library(ggplot2)
library(gganimate)
library(knitr)
library(kableExtra)
library(readr)
library(lubridate)
```

## Visualize with ggplot2

ggplot2 is a package for visualizing data in R. It's a "grammar of graphics".
With ggplot2, one can create all manner of visualizations. We can also explore gganimate, a package for bringing a visual to life. 

The Grammar of Graphics offers up *three* key elements when creating a visualization:
* data, or the dataset to be used in the visualization
* geom, or what type of visualization will be created
* aes, or the details of mapping the geom to the data, plus other aesthetics

The structure:

ggplot( *data to be used* ) +
  *geom to be used* ( mapping = *aes( like what is x, y, etc)* )

### Let's practice with some Trees data.
Reading in the Trees dataset from Edmonton's Open Data site

```{r edmonton trees}
trees_url <- "https://data.edmonton.ca/resource/93cp-z7sw.json?$limit=100000"  ## Just 100,000 records to make things a little faster
trees <- read.socrata(trees_url) %>% mutate(condition_percent = as.numeric(condition_percent),
                                            diameter_breast_height = as.numeric(diameter_breast_height))
head(trees)
```

### Here's an Example: Scatterplot
The *data* is trees. 
The *geom* is geom_point, for a scatterplot. 
The *aes* is what columns we assign to x and y. 

Run the Code chunk below to see the chart. 

```{r}
ggplot(trees) +
  geom_point(mapping = aes(x = diameter_breast_height, y = condition_percent))
```


### Tree Size and Condition Percent by Genus
Here, we will use three additional arguments with our geom: 
*color* to visually distinguish based on another attribute in the data
*alpha* to control transparency of the points
*position* to modify how the points are plotted (right on top of each other? jittered a little?)

```{r trees chart}
ggplot(trees) +
  geom_point(mapping = aes(x = diameter_breast_height, y = condition_percent, color = genus), alpha = 0.5, position = "jitter")

```

### Try it Yourself!
In the code chunk below, create the same chart but with colors representing 'location_type'. Some structure has been provided to get you started, but not all: aes() and all of its arguments are missing.

Stuck? Copy everything after "mapping = " and before the last parenthesis from the example above, and change *genus* to *location_type.*

```{r try trees chart,eval=F}
ggplot(trees) +
  geom_point(mapping = aes(x = diameter_breast_height, y = condition_percent, color = location_type), alpha = 0.8, position = "jitter")

```


## Local Area Unemployment Statistics in California

### Visualize labor force, unemployment rate for areas in California
First, let's create some subsets of the data. 

```{r variables}
edd_url <- "https://data.edd.ca.gov/Labor-Force-and-Unemployment-Rates/Local-Area-Unemployment-Statistics-LAUS-/e6gw-gvii"
laus_data <- read.socrata(edd_url)

la_county <- laus_data %>% filter(`area_name` == "Los Angeles County" & `seasonally_adjusted_y_n` == "N")

la_msa <- laus_data %>% filter(area_name == "Los Angeles County" | area_name == "Orange County" & `seasonally_adjusted_y_n` == "N")

la_csa <- laus_data %>% filter(area_name == "Los Angeles County" | area_name == "Orange County" | area_name == "Ventura County" | area_name == "Riverside County" | area_name == "San Bernardino County" & `seasonally_adjusted_y_n` == "N")

la_metro <- laus_data %>% filter(area_type == "Metropolitan Area" & area_name == "Los Angeles-Long Beach-Glendale MD" & `seasonally_adjusted_y_n` == "N")

cat("Nice! We now have some filtered views of the unemployment data.")
```


## Let's take a look at Los Angeles County's unemployment rates over time. 
We will use a new geom: *geom_line*

```{r LAUS exploration LA}
ggplot(la_county) +
  geom_line(mapping = aes(x = date, y = `unemployment_rate`))

```



### Your Turn: Metropolitan Statistical Area Plot
Fill in the mapping arguments for geom_line. Recall the structure from previous examples. 
- x = date
- y = unemployment_rate
- color = area_name

```{r msa, eval=F}
ggplot(la_msa) +
  geom_line(mapping = aes(x = date, y = `unemployment_rate`, color = area_name))
```

### CSA Plot
Let's make sure other counties around LA don't have the cliff that Orange County does. Run the code chunk to check.

```{r csa}
ggplot(la_csa) +
  geom_line(aes(x = date, y = unemployment_rate, color = area_name))

```

## Faceting!
Faceting creates small plots of the same geom based on some discrete variable you provide. 

```{r}
ggplot(la_csa) +
  geom_line(aes(date,unemployment_rate)) +
  facet_wrap(~ area_name)
```

## ggplot and dplyr for Initial Exploration and Error Validation of Data

What's with that cliff in 2010 for Orange County?
We can spot check with further visualization and functions from the dplyr package. 

```{r the OC}
the_oc <- laus_data %>% filter(area_name == "Orange County")
check_the_oc <- the_oc %>% group_by(year) %>% summarise(rown = length(unique(date)))

ggplot(check_the_oc) +
  geom_col(mapping = aes(x = year, y = rown))
```

## Further refine a visualization in all sort of ways
- Mutliple geoms
- Axis choices
- Themes
- Annotations
- And much much more!

```{r uiwc}
uiwc_data <- read_csv("https://data.edd.ca.gov/api/views/qbrj-7mnz/rows.csv?accessType=DOWNLOAD&bom=true&format=true", 
                      col_types = cols(
                        `Area Type` = "c",
                        `Area Name` = "c",
                        `Filed week ended` = col_date(format = "%m/%d/%Y"),
                        `Initial Claims` = "n",
                        `Reflecting Week Ended` = col_date(format = "%m/%d/%Y"),
                        `Continued Claims` = "n",
                        `Covered Employment` = "n",
                        `Insured Unemployment Rate` = "d"
                      ))

ggplot(uiwc_data) +
    geom_line(mapping = aes(x = `Filed week ended`, y = `Initial Claims`)) +
    geom_line(mapping = aes(x = `Filed week ended`, y = `Continued Claims`)) +
    geom_smooth(mapping = aes(x = `Filed week ended`, y = `Initial Claims`)) +
    geom_smooth(mapping = aes(x = `Filed week ended`, y = `Continued Claims`)) +
    theme_bw() +
    labs(x = "Date", y = "Claims, Initial and Continued", title = "Unemployment Insurance Weekly Claims") +
    annotate(geom = "text", x = as_date("2000-12-01"), y = 125000, label = "Initial Claims") +
    annotate(geom = "text", x = as_date("2000-12-01"), y = 750000, label = "Continued Claims")

``` 



















