---
title: "Visualize Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(RSocrata)
library(dplyr)
library(ggplot2)
library(gganimate)
library(knitr)
library(kableExtra)
```

## Visualize with ggplot2

ggplot2 is a package for visualizing data in R. It's a "grammar of graphics".
With ggplot2, one can create all manner of visualizations. We can also explore gganimate, a package for bringing a visual to life. 
## Read in some Steller Sea Lion counts from NOAA Alaska Open Data

```{r steller sea lions data}
sea_lions_url <- "https://noaa-fisheries-afsc.data.socrata.com/Species/2016-Counts-of-Steller-Sea-Lions-Pups-Adults-and-J/cntk-5zqj"
rsocrata_data <- read.socrata(sea_lions_url)

cat("Let's take a look at a few of the rows.\n")
head(rsocrata_data, n = 5) %>%  kable(format = 'markdown', align = "c")
```

## And Group by Region

```{r region}
region_counts <- group_by(rsocrata_data, Region) %>% 
  summarise(Non.Pup = sum(Non.Pup, na.rm = TRUE),
            Pup = sum(Pup, na.rm = TRUE))
region_counts
```

## Visually Display the Number of Pups and Non-Pups by Site

The Grammar of Graphics offers up *three* key elements when creating a visualization:
* data, or the dataset to be used in the visualization
* geom, or what type of visualization will be created
* aes, or the details of mapping the geom to the data, plus other aesthetics

The structure:

ggplot( *data to be used* ) +
  *geom to be used* ( mapping = *aes( like what is x, y, etc)* )
  
Run the code chunk below to see an example:

```{r seals chart, echo=TRUE}
ggplot(data = region_counts) +
  geom_col(mapping = aes(x = Region, y = Non.Pup))
```

## Try it on your own:
Create the same type of chart, but showing the count of *Pups* on the y-axis. Some structure has been provided. 

```{r try seals chart, eval=F}
ggplot() +
  geom_bar(mapping = aes())

```

## Trees! (in Edmonton) Examples

Reading in the Trees from Edmonton's Open Data site

```{r edmonton trees}
trees_url <- "https://data.edmonton.ca/resource/93cp-z7sw.json?$limit=100000"  ## Just 100,000 records to make things a little faster
trees <- read.socrata(trees_url) %>% mutate(condition_percent = as.numeric(condition_percent),
                                            diameter_breast_height = as.numeric(diameter_breast_height))
```

## Tree Size and Condition Percent by Genus
Here, we will use three additional arguments with our geom: 
*color* to visually distinguish based on another attribute in the data
*alpha* to control transparency of the points
*position* to modify how the points are plotted (right on top of each other? jittered a little?)

```{r trees chart}
ggplot(trees) +
  geom_point(mapping = aes(x = diameter_breast_height, y = condition_percent, color = genus),alpha = 0.5,position = "jitter")
```

In the code chunk below, create the same chart but with colors representing 'location_type'. Some structure has been provided to get you started.

```{r try trees chart,eval=F}
ggplot(trees) +
  geom_point(mapping = aes())
```


## Local Area Unemployment Statistics in California
### Visualize labor force, unemployment rate for areas in California

Let's take a look at Los Angeles County's unemployment rates over time. 
We will use a new geom: *geom_line*

```{r LAUS exploration LA}
laus_la <- laus_data %>% filter(`area_name` == "Los Angeles County" & `seasonally_adjusted_y_n` == "Y")
la_unem_rate <- ggplot(laus_la) +
                  geom_line(mapping = aes(x = date, y = `unemployment_rate`)) 
la_unem_rate
```



```{r facet laus}
ggplot(laus_msa) +
  geom_line(mapping = aes(x = date, y = `unemployment_rate`, color = area_name),show.legend = FALSE)
```

```{r LAUS boxplot}
ggplot(laus_la) +
  geom_boxplot(mapping = aes(x = month, y = unemployment_rate))
```


















