---
title: "Visualize Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(RSocrata)
library(dplyr)
library(ggplot2)
library(gganimate)
library(knitr)
library(kableExtra)
```

## Visualize with ggplot2

ggplot2 is a package for visualizing data in R. Think of it as a "grammar of graphics".
With ggplot2, one can create all manner of visualizations. We we also explore gganimate, a package for bringing a visual to life. 

## Read in some Steller Sea Lion counts from NOAA Alaska Open Data

```{r steller sea lions data}
url <- "https://noaa-fisheries-afsc.data.socrata.com/Species/2016-Counts-of-Steller-Sea-Lions-Pups-Adults-and-J/cntk-5zqj"
rsocrata_data <- read.socrata(url)

cat("Let's take a look at a subset of the rows.\n")
head(rsocrata_data, n = 5) %>%  kable(format = 'markdown', align = "c")
```

## And Group by Region

```{r region}
region_counts <- group_by(rsocrata_data, Region) %>% 
  summarise(Non.Pup = sum(Non.Pup, na.rm = TRUE),
            Pup = sum(Pup, na.rm = TRUE))
region_counts
```

## Visually Display the Number of Pups and Non-Pups by Site

The Grammar of Graphics offers up *three* key elements when creating a visualization:
* data, or what will be visualized
* geom, or how to display the data
* aes, or the aesthetics of the display

The structure:

ggplot( *data to be used* ) +
  *geom to be used* ( mapping = *aes( to be used like what is x, y, etc)* )

```{r ggplot, echo=TRUE}
ggplot(data = region_counts) +
  geom_col(mapping = aes(x = Region, y = Non.Pup))
```

## Try it on your own:
Create the same chart, but showing the count of *Pups* on the y-axis. 

```{r try grammar, eval=F}


```

## Trees! (in Edmonton)

Reading in the Trees from Edmonton's Open Data site ()

```{r edmonton trees}
url <- "https://data.edmonton.ca/resource/93cp-z7sw.json?$limit=10000"  ## Just 10,000 records to make things a little faster
trees <- read.socrata(url)
```

## Are larger trees in better or worse condition?

```{r}
ggplot(trees) +
  geom_point(aes(diameter_breast_height,condition_percent))
```


## Is a certain genus of tree suffering more than others?

```{r}
ggplot(trees) +
  geom_count(aes(genus,condition_percent))
```





















